### 2017-06-19 sections/redsvd.tex

\subsection{RedSVD}

The exact SVD of a $m \times n$ matrix has computational complexity \newline $O(min(mn^2, m^2n))$ using the "big-O" notation. When applied on large data sets it tends to be very time-consuming. In \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.

As the implementation of the randomized SVD algorithm RedSVD library was used. RedSVD is an open-source software package for solving several matrix decompositions. The particular implementation of randomized version of decomposition is based on the algorithm described in \cite{Halko2011}. However, although the algorithm in RedSVD samples both rows and columns, the original algorithm samples in one way. The brief summary of the algorithm steps follows below.

% redsvd algorithm description
Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$ is chosen, then the result is full SVD). As the first step, a random Gaussian matrix $\mtrx{O}$ with $m$ rows and $r$ columns is created and a sample matrix $\mtrx{Y} = \mtrx{A}^{\mathsf{T}} \mtrx{O}$ is computed. Then the Gram-Schmidt process is applied to $\mtrx{Y}$ so that each column of $\mtrx{Y}$ is orthonormalized. Then matrix $\mtrx{B} = \mtrx{A}\mtrx{Y}$ with $n$ rows and $r$ columns is computed. Although the size of $\mtrx{Y}$ is much smaller than that of $\mtrx{A}$, $\mtrx{Y}$ holds the information of $\mtrx{A}$, i.e. $\mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} \approx \mtrx{A}$. Intuitively, the row information is compressed by $\mtrx{Y}$ and can be decompressed by $\mtrx{Y}^{\mathsf{T}}$.

Similarly, another random Gaussian matrix $\mtrx{P}$ with $r$ rows and $r$ columns is created, and matrix $\mtrx{Z} = \mtrx{B}\mtrx{P}$ is computed. As in the previous case, the columns of $\mtrx{Z}$ are orthonormalized by the Gram-Schmidt process and $\mtrx{Z}\mtrx{Z}^{\mathsf{T}} \mtrx{B} \approx \mtrx{B}$. Then matrix $\mtrx{C} = \mtrx{Z}^{\mathsf{T}} \mtrx{B}$ is computed.

Finally, the algorithm performs SVD of $\mtrx{C}$ using traditional exact SVD solver and obtains $\mtrx{C} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$, where $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ are orthonormal matrices, and $\mtrx{\widetilde{S}}$ is a diagonal matrix whose entries are singular values of matrix $\mtrx{C}$. Since matrix $\mtrx{C}$ is usually very small, computational time of this step is negligible.

$\mtrx{A}$ is then decomposed as

\begin{eqnarray}
\mtrx{A} &\approx& \mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} =
\\
&=& \mtrx{B}\mtrx{Y}^{\mathsf{T}} \approx \nonumber
\\
&\approx& \mtrx{Z}\mtrx{Z}^{\mathsf{T}}\mtrx{B}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{C}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} \mtrx{\widetilde{V}}^{\mathsf{T}}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} (\mtrx{Y}\mtrx{\widetilde{V}})^{\mathsf{T}}. \nonumber
\label{eq:redsvd}
\end{eqnarray}

\noindent
Both $\mtrx{Z}\mtrx{\widetilde{U}}$ and $\mtrx{Y}\mtrx{\widetilde{V}}$ are orthornormal, $\mtrx{Z}\mtrx{\widetilde{U}}$ contains approximations of left singular vectors, and $\mtrx{Y}\mtrx{\widetilde{V}}$ contains approximations of right singular vectors of matrix $\mtrx{A}$. $\mtrx{\widetilde{S}}$ contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.

% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
%Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation. Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$. Note that $\mtrx{AA^T}$ is the dataset covariance matrix. Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$. Statistically, the SVD of $\mtrx{BB^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.