### 2017-06-19 sections/randomized-svd.tex
Update on Overleaf.

\subsection{Randomized SVD}

The exact SVD of a $m \times n$ matrix has computational complexity \newline $\mathrm{O}(\mathrm{min}(mn^2, m^2n))$ using the "big-O" notation. When applied on large data sets it tends to be very time-consuming. In \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.

The particular implementation of the randomized decomposition is based on the algorithm described in \cite{Halko2011}. The authors proposed an algorithm for efficient computation of low-rank approximation to a given matrix. The algorithm can be split into two main computational stages.

The first stage is to construct a low-dimensional subspace that captures the action of the matrix. To be more formal, this stage is to compute an approximate basis for the range of the input matrix $\mtrx{A}$. This basis matrix $\mtrx{Q}$ is required to have orthonormal columns and

\begin{equation}
\mtrx{A} \approx \mtrx{Q} \mtrx{Q}^{\mathsf{T}} \mtrx{A}.
\end{equation}

\noindent
Matrix $\mtrx{Q}$ is desired to contain as few columns as possible while producing accurate approximation of matrix $\mtrx{A}$ at the same time.

The second stage is to use $\mtrx{Q}$ to obtain approximate SVD factorization of $\mtrx{A}$. This can be achieved using simple deterministic steps:

\begin{enumerate}
\item Construct $\mtrx{B} = \mtrx{Q}^{\mathsf{T}} \mtrx{A}$.
\item Compute an exact SVD of the small matrix: $\mtrx{B}=\mtrx{W}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$.
\item Set $\mtrx{\widetilde{U}}=\mtrx{Q}\mtrx{W}$.
\end{enumerate}

The main challenge is therefore to efficiently construct $r$ orthonormal vectors forming the matrix $\mtrx{Q}$ that (nearly) span the range of $\mtrx{A}$; $r$ is the desired rank of approximation and is supposed to be substantially less then both dimensions of $\mtrx{A}$. After that an SVD that closely approximates $\mtrx{A}$ can be constructed (closely in the sense that the spectral norm of the difference between $\mtrx{A}$ and the approximation to $\mtrx{A}$ is small relative to the spectral norm of $\mtrx{A}$).

In order to estimate the range of matrix $\mtrx{A}$, it is applied to a collection of $r$ random vectors. The result of applying $\mtrx{A}$ to any vector is a vector in the range of $\mtrx{A}$, and if the matrix is applied to $r$ random vectors, the results will nearly span the range of $\mtrx{A}$ with extremely high probability. Mathematical proofs given in \cite{Halko2011} and \cite{Witten2015} show that the probability of missing a substantial part of the range of $\mtrx{A}$ is negligible if the vectors to which we apply $\mtrx{A}$ are sufficiently random (i.e. entries of these vectors are independent and identically distributed).

Therefore, the matrix $\mtrx{A}$ is applied to a random Gaussian matrix $\mtrx{\Omega}$ that contains $r$ columns with random normally distributed entries yielding the matrix $\mtrx{Y} = \mtrx{A} \mtrx{\Omega}$. Applying the Gram-Schmidt process (or any other method for constructing QR decomposition) produces the decomposition $\mtrx{Y}=\mtrx{Q}\mtrx{R}$, where columns of $\mtrx{Q}$ are an orthonormal basis for the range of $\mtrx{Y}$, and since columns of $\mtrx{Y}$ nearly span the range of $\mtrx{A}$, $\mtrx{Q}$ is an orthonormal basis for the approximate range of $\mtrx{A}$.

$\mtrx{A}$ is then decomposed as
\begin{eqnarray}
\mtrx{A} &\approx& \mtrx{Q}\mtrx{Q}^{\mathsf{T}}\mtrx{A} =
\\
&=& \mtrx{Q}\mtrx{B} = \nonumber
\\
&=& \mtrx{Q}\mtrx{W}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}. \nonumber
\end{eqnarray}

\noindent
The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of left and right singular vectors of matrix $\mtrx{A}$, and a nonnegative, diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$. For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{k})$ floating-point operations, substantially less than classical algorithms.